---
title: "Data 607 - Acquisition and Management - Assignment 1"
author: "Enid Roman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
```

# In 2020 there was a high level of enthusiasm for voting in the general election. In the beginning of the year, voter registration had increased. Then COVID-19 pandemic hit and new registrations had suddenly decreased.

# As per the data taken registration numbers were even stronger in early 2020 than early 2016. Things changed dramatically in March, at least in the 12 states that data were taken.

# In this analysis we will focus on the comparison in Voter Registration in 2016 vs 2020. We will see the difference in numbers of Voter Registration.  

# This data was taken from the article Voter Registration Are Way, Way Down During The Pandemic, June 20, 2020 by Kaleigh Rogers and Nathaniel Rakich. https://fivethirtyeight.com/features/voter-registrations-are-way-way-down-during-the-pandemic/

# This data was collected for https://electioninnovation.org/wp-content/uploads/2020/06/New_Voter_Registrations.pdf

# The data and attached information was taken from https://github.com/fivethirtyeight/data/tree/master/voter-registration

# The data contains total # of people registered to vote from early 2016 vs early 2020 from January to April in 11 states and to May in Washington, DC. Data was not available for the other states.

# The data is broken down into four variables: Jurisdiction, Year, Month, New.registered.voters.

```{r}
# I uploaded the libraries I might need. 

library("pacman")
library("tidyverse")
library("dplyr")
library("ggplot2")
library("scales")
```

# I uploaded the csv dataset to my github where I created a repository for the dataset. I then uploaded the dataset to RStudio.

```{r}
# I uploaded the csv dataset to my github where I created a repository for the dataset. I then uploaded the dataset to RStudio.

theUrl <- "https://raw.githubusercontent.com/enidroman/data607_acquisition_and_managent/main/new-voter-registrations.csv"

voter_registration <- read.table(file=theUrl, header=TRUE, sep=",")
voter_registration
```
# The data frame voter_registration has 106 observations and 4 variables.
# The variables type, Jurisdiction, and Month are in character mode and Year and New.registered.voters that are integers.

```{r}
# Checked the structure of the data with the function str():
str(voter_registration, vec.len = 1)
```

# In this dataset "Year" is actually a categorial data not an integer. 

```{r}
# I changed "Year" from integer to a factor.

voter_registration$Year <- as.factor(voter_registration$Year)
str(voter_registration, vec.len = 1)
```


# The data is 106 lines in length and the "Year" are 2016 and 2020. The minimum of New Registered Voters is 589, Median is 33302, Mean is 48223, and Max New Registered Voters is 238381. 

```{r}
#I did a summary of the dataset.

summary(voter_registration)
```

# Since data was not available for all of states for the month of May, I removed the rows that contain the month of May from the data frame.

#I tried using the below code to remove the rows containing the month of May but I was not successful.

```{r}
#voter_registration_to_remove <- subset(voter_registration, "Month" != "May") 
```

#I then went through the dataset and note down the index # of the rows that contained the month of May and I was able to remove those rows by the index #. 

```{r}
voter_registration_to_remain <- voter_registration[-c(37, 42, 71, 76, 81, 86, 91, 96, 101, 106),]
voter_registration_to_remain
```
#Since I removed the rows that contained the month of May, I reset the index.

```{r}
voter_registration_new <- voter_registration_to_remain                           
rownames(voter_registration_new) <- 1:nrow(voter_registration_new)    
voter_registration_new                                   
```
# In order to compare each the Registered Voters for the month of Jan, Feb, Mar and April of each year, I made 2 columns with the total of Registered Voters for the Year 2016 and Year 2020. 

```{r}
voter_registration_new_wide <- pivot_wider(voter_registration_new, names_from = Year, values_from = New.registered.voters)
voter_registration_new_wide
```

```{r}
#I renamed the columns.

colnames(voter_registration_new_wide)[1:4] = c("STATE", "MONTH", "REGISTERED_VOTERS_2016", "REGISTERED_VOTERS_2020")
voter_registration_new_wide
```
```{r}
# I then created a column for the difference of Registgered Voters in 2016 and 2019.

voter_registration_new_wide$REGISTERED_VOTERS_DIFF <- (voter_registration_new_wide$REGISTERED_VOTERS_2016 - voter_registration_new_wide$REGISTERED_VOTERS_2020)
voter_registration_new_wide
```
# Now the dataset is 48 rows in length. You can also see the Min, Median, Mean, Max for Registered Voters 2016, 2020, and the Regisgtered Voters Difference for the 12 States. For the Year 2016 the Min was 2840, the Median was 40737, the Mean was 54068, and the Max was 185478. For the Year 2020 the Min was 589, the Median was 31682, the Mean was 47192, and the Max was 238281. For the difference between the 2 years the Min was -134904, the Median was 3347 the Mean was 6877, and the Max was 146508. 

# With this summary you can already see that there is a big difference in drop of Registered Voters for the year 2020 in the 12 States. Even though we see that the Max of Registered Voters was in 2020 with a 238281 Registered Voters. This needs further investigation.
```{r}
#I again did a new summary. 

summary(voter_registration_new_wide)
```

# I did a ggplot geometric line of the Registered Votes for each State per year.

```{r}
# Viewing the unique values in the "Month" variable===
unique(voter_registration_new_wide$MONTH)

# Identifying the class of the "Month" variable===
class(voter_registration_new_wide$MONTH)
```

```{r}
# In order for the Months to be in sequential order in the X axes I had to convert the Month from Character to Factor and set the order. 

voter_registration_new_wide <- voter_registration_new_wide %>%
  mutate(MONTH = factor(MONTH, levels = c("Jan","Feb","Mar","Apr")))
```

# In this graph you can see in 2016 that the Total Registered Voters was increasing for the month of Jan and Feb for most states accept for Illinois, District of Columbia, Florida, and North Carolina, and Arizona. The only States that increased from Mar to Apr was Colorado, North Carolina, and Illinois. Delaware seem to maintain the same through all 4 months. 

```{r}
# I did a ggplot geometric line to compare the Registered Voters for each of the 12 State for 2016. 

# I did notice that I would need to use different colors that is not colour because District of Columbia and Florida are the same shade of green. 

ggplot(voter_registration_new_wide, aes(x = MONTH, y = REGISTERED_VOTERS_2016, colour = STATE, group = STATE)) +
  geom_line() +
   geom_point()
```

# In this graph you can see in 2020 that the Total Registered Voters had increased somewhat for the month of Jan to Feb accept for Texas that decreased slightly and North Carolina dropped significantly. California had increased significantly. Feb to April they all decreased significantly accept for Texas, North Carolina, Virginia and Maryland had a slight borderline from Feb to Mar. Again Delaware maintained the same thought the 4 months. 

```{r}
# I did a ggplot geometric line to compare the Registered Voters for each of the 12 State for 2020

ggplot(voter_registration_new_wide, aes(x = MONTH, y = REGISTERED_VOTERS_2020, colour = STATE, group = STATE)) +
  geom_line() +
   geom_point()
```

# In this graph you can see that Registered Voters difference were closed to or below 0 from Jan to Feb. California tremedously decreased from Jan to Feb. Feb to Mar all State increased slighly accept California that increased significantly. Mar to Apr all States increased a little accept for Texas that increased more and California increased significantly. 

```{r}
# I did a ggplot geometric line to compare the Registered Voters difference for each of the 12 State for 2016 vs 2020

ggplot(voter_registration_new_wide, aes(x = MONTH, y = REGISTERED_VOTERS_DIFF, colour = STATE, group = STATE)) +
  geom_line() +
   geom_point()

```

# It is clear that despite voters strong interest in the 2020 election, the coronavirus had already made it harder for new voters to participate in it.

# Based on these graphs and data further invesitigation needs to be done on the reason for the sudden increases and decreases in California. 

# Below I compare total 12 states for Jan, Feb, Mar, and April each year.

```{r}
# Total all 12 states for Jan, Feb, Mar, and April in 2016.

voter_registration_new_2016 <-voter_registration_new_wide %>%
  group_by(MONTH) %>%
  summarise(REGISTERED_VOTERS_2016 = sum(REGISTERED_VOTERS_2016)) 
voter_registration_new_2016
  
```
# Here you can see the tremedous increase and slight decrease for all 12 states from Jan to April in 2016.

```{r}
ggplot(data=voter_registration_new_2016, aes(x=MONTH, y=REGISTERED_VOTERS_2016, group=1)) +
  geom_line()+
  geom_point()
```

```{r}
# Total all 12 states for Jan, Feb, Mar, and April in 2020.

voter_registration_new_2020 <-voter_registration_new_wide %>%
  group_by(MONTH) %>%
  summarise(REGISTERED_VOTERS_2020 = sum(REGISTERED_VOTERS_2020)) 
voter_registration_new_2020
```

# Here you can see the tremedous drop for all 12 states from Jan to April in 2020.

```{r}
ggplot(data=voter_registration_new_2020, aes(x=MONTH, y=REGISTERED_VOTERS_2020, group=1)) +
  geom_line()+
  geom_point()
```

# It is clear that despite voters strong interest in the 2020 election, the coronavirus had already made it harder for new voters to participate in it.

# Based on these graphs and data further invesitigation needs to be done on the reason for the sudden increases and decreases in California. 