---
title: "Movie_Rating with SQL"
author: "Enid Roman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I chose six recent films and asked five different imaginary individuals to rate each of the movies they had seen from a scale of 1 to 5. 

Data Collection

I created a database called movie_rating in a MySQL workbench to store the data in individual tables:

Table 1: movie
Table 2: name
Table 3: review

The SQL code for table creation:

CREATE TABLE ‘movie’ (
 ‘movie_id’ int NOT NULL,
 ‘title’ varchar(45) NOT NULL,
 ‘length’ varchar(45) NOT NULL,
PRIMARY KEY (‘movie_id’))

CREATE TABLE ‘name’(
  ‘name_id’ int NOT NULL,
  ‘first_name’ varchar(45) NOT NULL,
  ‘age’ varchar(45) NOT NULL,
PRIMARY KEY (‘name_id))

CREATE TABLE ‘review’ (
  ‘review_id’ int NOT NULL,
  ‘movie_id’ varchar(45) NOT NULL,
  ‘name_id’ varchar(45) NOT NULL,
  ‘rating’ int DEFAULT NULL,
  ‘review’ varchar(45) DEFAULT NULL,
PRIMARY KEY (‘review_id’)
)


First I installed and uploaded the packages I needed. 

```{r}
# First I installed and uploaded the packages I needed. 

#install.packages("RMySQL")
#install.packages("DBI")
library(RMySQL)
library(DBI)
library(dbplyr)
library(tidyverse)
library(ggplot2)
```

You need to create a password to access the local database. 

```{r}
# I then connected MySQL to R to upload my database, movie_rating

mydb = dbConnect(RMySQL::MySQL(),
       dbname='movie_rating',
       host='127.0.0.1',
       port=3306,
       user='root',
       password=rstudioapi::askForPassword("database password"))
```

```{r}
# I previewed the tables.

dbListTables(mydb)
```
Here I see the oldest of the individual is 39 and the youngest is 7.

```{r}
# I wrote a query to show name table.

name_table <- dbSendQuery(mydb, "SELECT * FROM name;")
dbFetch(name_table)
```

Here I see the longest in length time is The Good Nurse and the shortest is Texas Chainsaw Massacre. 

```{r}
# I wrote a query to show movie table. 

movie_table <- dbSendQuery(mydb, "SELECT * FROM movie;")
dbFetch(movie_table)
```
```{r}
# I wrote a query to show review table.

review_table <- dbSendQuery(mydb, "SELECT * FROM review")
dbFetch(review_table)
```


```{r}
# I then joined all three tables movie, name, and review to create one table called movie_rating.

movie_rating <- dbSendQuery(mydb, "SELECT 
M.title As 'Title',
M.length AS 'Length',
N.first_name AS 'Name',
N.age AS 'Age',
R.rating As 'Rating',
R.review AS 'Review'
FROM movie AS M
JOIN review AS R
ON M.movie_id = R.movie_id
JOIN name AS N
ON N.name_id = R.name_id;")
#dbFetch(movie_rating)
data<-fetch(movie_rating)
print(data)
```


```{r}
# Checked the structure of the data. 30 rows. 6 columns. 

str(data, vec.len = 1)
```

Here I see The Adam Project had the highest average rating with 4.8 and Tall Girl 2 has the lowest with 2.8.

```{r}
# I did a group by to see the average score for each movie rated. 

new_data <- data %>%
  filter(!is.na(Rating)) %>%
   group_by(Title) %>%
    summarise(Avg_Score = mean(as.numeric(Rating))) %>%
      arrange(desc(Avg_Score))
        new_data
```

Same results as the above.

```{r}
# I did a ggplot of the Average Score for each movie for visualization purpose. 

new_data %>%
ggplot +
geom_col(aes(Avg_Score, Title))
```

The Adam Project has the most 5 rating with 4 counts. Texas Mascare had 0 rating from a 7 year old who thought it was too bloody.

The Good Nurse got 1 - 2 Rating and 1 - 3 Rating and 3 - 4 Rating. That is about average rating.

The Adam Project got 1 - 3 Rating and 3 - 4 Rating. That is above average, 1st in place. 

Texas Chainsaw Massacre got 1 - 0 Rating, 2 - 4 Rating, 2 - 5 Rating.

Tall Girl got 3 - 2 Rating and 2 - 4 Rating. That is below average. Ranked the lowest. 

Pinocchio got 1 - 3 Rating, 1 - 4 Rating, and 3 - 5 Rating. This is above average, 2nd in place. 

Against The Ice got 1 - 2 Rating, 1 - 3 Rating, and 3 - 5 Rating. This is average. 

```{r}
# I did another group by to see the count for each rating per movie. 

count_data <- data 
count_data %>% group_by(Title, Rating) %>% summarise(count = n())%>%
arrange(desc(Title))
```

In conlusion more research has to be done on the ratings for these 5 movies to see what motivated these individuals to give them the rating they received. For now we can go as per the written reviews section for the reason of their ratings. For example Texas Chainsaw was bloody. Some gave it high ranking for that because the like horrow movies and some gave 0 because it was too bloody for her. We might want to look at the length of the movies. For example, The Good Nurse has a longest length of 121 minutes out of the 6 movies. Texas Chainsaw Massacre was the shortest with 81 minutes. The length of the movie could also affect the rating of the movies. Also we might want to look at the age of the individuals. The age could affect the rating of the movies. We would also need to take a take rating from a larger group to get a broader analysis. 

For now The Adam Project ranks # 1 and Tall Girl ranks # 6 as per the 5 individuals. 

























